---
title: "Histone Acetyl RNA-seq Enrichment analysis"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

Source: https://github.com/markziemann/histone_acetyl_rnaseq

# Introduction

In this study we are looking at the effect of knock-down of two genes which are required
for the supply of acetyl units.
It is hypothesised that knock-down of these genes could lead to problems with
the response to exercise.
We have both ctrl and kd under either sedentary or exercise conditions.

The focus of this report is to perform mitch enrichment analysis of the DE tables
I have already generated.

Here are the DE analyses I've already performed.

| Contrast number | DE file | Contrast |
| --- | --- | --- |
| 1 | dge1pkd.tsv | Effect of ATP-CL KD in Sed animals |
| 2 | dge2pkd.tsv | Effect of ATP-CL KD in Ex animals |
| 3 | dge3pkd.tsv | Effect of ACSS2 KD in Sed animals |
| 4 | dge4pkd.tsv | Effect of ACSS2 KD in Ex animals |
| 5 | dge5.tsv | Effect of Sed vs Ex (ATP-CL gene ctrl) |
| 6 | dge6kd.tsv | Effect of Sed vs Ex (ATP-CL gene KO) |
| 7 | dge7.tsv | Effect of Sed vs Ex (ACSS2 gene ctrl) |
| 8 | dge8kd.tsv | Effect of Sed vs Ex (ACSS2 gene kd) |

First I will run mitch with each contrast separately,
then joint enrichment in the following ways:

* #1 and #2 effect of ATP-CL KD in Sed and Ex animals

* #3 and #4 effect of ACSS2 KD in Sed and Ex animals

* #5 and #6 Effect of Sed vs Ex in ctrl and ATP-CL gene KD

* #7 and #8 Effect of Sed vs Ex in ctrl and ACSS2 gene KD

Finally all contrasts altogether.

```{r,packages}

suppressPackageStartupMessages({
    library("plyr")
    library("zoo")
    library("tidyverse")
    library("reshape2")
    library("DESeq2")
    library("gplots")
    library("fgsea")
    library("MASS")
    library("mitch")
    library("eulerr")
    library("limma")
    library("topconfects")
    library("kableExtra")
    library("vioplot")
    library("beeswarm")
})

```

## Import DE data

```{r,load1}

de1 <- read.csv("dge1pkd.tsv",header=TRUE,row.names=1,sep="\t")
de2 <- read.csv("dge2pkd.tsv",header=TRUE,row.names=1,sep="\t")
de3 <- read.csv("dge3pkd.tsv",header=TRUE,row.names=1,sep="\t")
de4 <- read.csv("dge4pkd.tsv",header=TRUE,row.names=1,sep="\t")
de5 <- read.csv("dge5.tsv",header=TRUE,row.names=1,sep="\t")
de6 <- read.csv("dge6kd.tsv",header=TRUE,row.names=1,sep="\t")
de7 <- read.csv("dge7.tsv",header=TRUE,row.names=1,sep="\t")
de8 <- read.csv("dge8kd.tsv",header=TRUE,row.names=1,sep="\t")

```

## Unidimensional mitch

First format the data into a friendly format.

```{r,1d}

genesets <- gmt_import("https://ziemann-lab.net/public/msigdb_mouse/reactome.v5.2.symbols_mouse.gmt")

de1$gene <- sapply(strsplit(rownames(de1)," "),"[[",2)
de1 <- aggregate(stat ~ gene,de1,mean)
de1rn <- de1
rownames(de1) <- de1$gene
de1$gene=NULL

de2$gene <- sapply(strsplit(rownames(de2)," "),"[[",2)
de2 <- aggregate(stat ~ gene,de2,mean)
de2rn <- de2
rownames(de2) <- de2$gene
de2$gene=NULL

de3$gene <- sapply(strsplit(rownames(de3)," "),"[[",2)
de3 <- aggregate(stat ~ gene,de3,mean)
de3rn <- de3
rownames(de3) <- de3$gene
de3$gene=NULL

de4$gene <- sapply(strsplit(rownames(de4)," "),"[[",2)
de4 <- aggregate(stat ~ gene,de4,mean)
de4rn <- de4
rownames(de4) <- de4$gene
de4$gene=NULL

de5$gene <- sapply(strsplit(rownames(de5)," "),"[[",2)
de5 <- aggregate(stat ~ gene,de5,mean)
de5rn <- de5
rownames(de5) <- de5$gene
de5$gene=NULL

de6$gene <- sapply(strsplit(rownames(de6)," "),"[[",2)
de6 <- aggregate(stat ~ gene,de6,mean)
de6rn <- de6
rownames(de6) <- de6$gene
de6$gene=NULL

de7$gene <- sapply(strsplit(rownames(de7)," "),"[[",2)
de7 <- aggregate(stat ~ gene,de7,mean)
de7rn <- de7
rownames(de7) <- de7$gene
de7$gene=NULL

de8$gene <- sapply(strsplit(rownames(de8)," "),"[[",2)
de8 <- aggregate(stat ~ gene,de8,mean)
de8rn <- de8
rownames(de8) <- de8$gene
de8$gene=NULL

```
Enrich it.

```{r,enrich1}

res1 <- mitch_calc(de1, genesets, priority="effect",cores=8)
head(res1$enrichment_result,20) %>% kbl(caption="DE1: Reactome") %>% kable_styling()
mitch_report(res=res1,outfile="mitch_de1.html",overwrite=TRUE)

res2 <- mitch_calc(de2, genesets, priority="effect",cores=8)
head(res2$enrichment_result,20) %>% kbl(caption="DE2: Reactome") %>% kable_styling()
mitch_report(res=res2,outfile="mitch_de2.html",overwrite=TRUE)

res3 <- mitch_calc(de3, genesets, priority="effect",cores=8)
head(res3$enrichment_result,20) %>% kbl(caption="DE3: Reactome") %>% kable_styling()
mitch_report(res=res3,outfile="mitch_de3.html",overwrite=TRUE)

res4 <- mitch_calc(de4, genesets, priority="effect",cores=8)
head(res4$enrichment_result,20) %>% kbl(caption="DE4: Reactome") %>% kable_styling()
mitch_report(res=res4,outfile="mitch_de4.html",overwrite=TRUE)

res5 <- mitch_calc(de5, genesets, priority="effect",cores=8)
head(res5$enrichment_result,20) %>% kbl(caption="DE5: Reactome") %>% kable_styling()
mitch_report(res=res5,outfile="mitch_de5.html",overwrite=TRUE)

res6 <- mitch_calc(de6, genesets, priority="effect",cores=8)
head(res6$enrichment_result,20) %>% kbl(caption="DE6: Reactome") %>% kable_styling()
mitch_report(res=res6,outfile="mitch_de6.html",overwrite=TRUE)

res7 <- mitch_calc(de7, genesets, priority="effect",cores=8)
head(res7$enrichment_result,20) %>% kbl(caption="DE7: Reactome") %>% kable_styling()
mitch_report(res=res7,outfile="mitch_de7.html",overwrite=TRUE)

res8 <- mitch_calc(de8, genesets, priority="effect",cores=8)
head(res8$enrichment_result,20) %>% kbl(caption="DE8: Reactome") %>% kable_styling()
mitch_report(res=res8,outfile="mitch_de8.html",overwrite=TRUE)

```

## Bidimensional

B1 (DE1 and DE2): effect of ATP-CL KD in Sed and Ex animals

```{r,bi1}

b1 <- merge(de1,de2,by=0)
rownames(b1) <- b1$Row.names
b1$Row.names=NULL
colnames(b1) <- c("Sed","Ex")
resb1 <- mitch_calc(b1, genesets, priority="effect",cores=8)
head(resb1$enrichment_result,20) %>% kbl(caption="effect of ATP-CL KD in Sed and Ex animals") %>% kable_styling()
mitch_report(res=resb1,outfile="mitch_b1.html",overwrite=TRUE)

```

B2 (DE3 and DE4): effect of ACSS2 KD in Sed and Ex animals

```{r,bi2}

b2 <- merge(de3,de4,by=0)
rownames(b2) <- b2$Row.names
b2$Row.names=NULL
colnames(b2) <- c("Sed","Ex")
resb2 <- mitch_calc(b2, genesets, priority="effect",cores=8)
head(resb2$enrichment_result,20) %>% kbl(caption="effect of ACSS2 KD in Sed and Ex animals") %>% kable_styling()
mitch_report(res=resb2,outfile="mitch_b2.html",overwrite=TRUE)

```

B3 (DE5 and DE6): Effect of Sed vs Ex in ctrl and ATP-CL gene KD

```{r,bi3}

b3 <- merge(de5,de6,by=0)
rownames(b3) <- b3$Row.names
b3$Row.names=NULL
colnames(b3) <- c("ctrl","AcyKD")
resb3 <- mitch_calc(b3, genesets, priority="effect",cores=8)
head(resb3$enrichment_result,20) %>% kbl(caption="effect of  Sed vs Ex in ctrl and ATP-CL gene KD") %>% kable_styling()
mitch_report(res=resb3,outfile="mitch_b3.html",overwrite=TRUE)

```

B4 (DE7 and DE8): Effect of Sed vs Ex in ctrl and ACSS2 gene KD

```{r,bi4}

b4 <- merge(de7,de8,by=0)
rownames(b4) <- b4$Row.names
b4$Row.names=NULL
colnames(b4) <- c("ctrl","Acss2KD")
resb4 <- mitch_calc(b4, genesets, priority="effect",cores=8)
head(resb4$enrichment_result,20) %>% kbl(caption="effect of  Sed vs Ex in ctrl and Acss2 gene KD") %>% kable_styling()
mitch_report(res=resb4,outfile="mitch_b4.html",overwrite=TRUE)

```

## Multi-dimensional

```{r,multi1}

colnames(de1rn) <- c("gene","de1")
colnames(de2rn) <- c("gene","de2")
colnames(de3rn) <- c("gene","de3")
colnames(de4rn) <- c("gene","de4")
colnames(de5rn) <- c("gene","de5")
colnames(de6rn) <- c("gene","de6")
colnames(de7rn) <- c("gene","de7")
colnames(de8rn) <- c("gene","de8")

ll <- list("de1"=de1rn,"de2"=de2rn,"de3"=de3rn,"de4"=de4rn,"de5"=de5rn,"de6"=de6rn,"de7"=de7rn,"de8"=de8rn)
jj <- join_all(ll)
head(jj)
rownames(jj) <- jj$gene
jj$gene=NULL
dim(jj)

resm <- mitch_calc(jj, genesets, priority="effect",cores=8)
head(resm$enrichment_result,20) %>% kbl(caption="Multimitch") %>% kable_styling()
mitch_report(res=resm,outfile="mitch_m.html",overwrite=TRUE)

```

## Session information

```{r,session}

sessionInfo()

```
